Fernando LOOCV

### 2016-11-22 reports/loocv_reduced.Rmd
Transfer prediction plots to report section

# Prediction For Reduced Data
```{r message = FALSE, echo = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
               "viridis", "tidyr", "modelr")

knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = "~/gamazon/",
                     inline = TRUE,
                     verbose = FALSE)
knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
                      warning = FALSE, echo = FALSE)
```

```{r Load-LOOCV-Data, cache = TRUE}
# Load log files and data
run_log <- fread("./data/derived/na_fraction_pred_log.txt")

# LOOCV 
loocv_nms <- run_log[CV == "LOOCV", Job_ID, ]
loocv <- lapply(seq_along(loocv_nms), FUN = function(i) {
  readRDS(paste0("./data/derived/predictions/", loocv_nms[i], ".RDS"))
})
loocv <- rbindlist(loocv)
```

```{r Rename-Traits, cache = TRUE}
loocv_clean <- loocv %>%
  mutate(Trait = fct_recode(Trait, 
                            "FAT" = "FETT",
                            "PRO" = "RPR",
                            "DMY" = "GTM",
                            "DMC" = "GTS",
                            "SUG" = "XZ"),
         Trait = fct_relevel(Trait, c("DMY", "DMC", "ADF", "FAT", "PRO", "STA",
                                      "SUG")),
         Dent_NA_Fraction = as.factor(as.character(Dent_NA_Fraction)),
         Flint_NA_Fraction = as.factor(as.character(Flint_NA_Fraction)),
         Residuals = y - yhat,
         Predictor = fct_recode(Predictor,
                                "mRNA" = "mrna",
                                "SNP" = "snp")
  )
```

## Predictive ability for non-imputed data
Peak at predictive abilities based on non-imputed predictor information and rank
the results by their predictive ability.
```{r Non-Imputed-Pred-Ability, cache = TRUE}
loocv_tbl_cap <- paste("Predictive abilities for the 'reduced' set of hybrids",
                       "using two different predictors and seven agronomic",
                       "traits. Results are ranked by predictive ability.")
loocv_clean %>%
  filter(Dent_NA_Fraction == 0, Flint_NA_Fraction == 0) %>%
  mutate(Predictor = as.factor(Predictor)) %>%
  droplevels() %>%
  group_by(Predictor, Trait) %>%
  summarize(Pred_Ability = round(cor(y, yhat), digits = 2)) %>%
  arrange(desc(Pred_Ability)) %>%
  rename("Predictive Ability" = Pred_Ability) %>%
  knitr::kable(caption = loocv_tbl_cap, booktabs = TRUE)
```

## Predicted vs. residuals
```{r Complete-Data-Residuals, cache = TRUE, dependson = "Rename-Traits"}
full_pred_resid <- loocv_clean %>%
  filter(Dent_NA_Fraction == 0, Flint_NA_Fraction == 0) %>%
  group_by(Predictor) %>%
  do(plot = ggplot(., aes(x = yhat, y = Residuals)) +
       geom_point(size = 0.5) +
       geom_ref_line(h = 0) +
       facet_wrap(~ Trait, nrow = 2, scales = "free") +
       theme(axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1)) +
       xlab("Predicted") +
       ylab("Residuals") +
       ggtitle(.$Predictor))
invisible(lapply(full_pred_resid$plot, FUN = print))
```

## Predicted vs. Observed
```{r Complete-Data-y-yHat, cache = TRUE, dependson = "Rename-Traits"}
full_y_yhat <- loocv_clean %>%
  filter(Dent_NA_Fraction == 0, Flint_NA_Fraction == 0) %>%
  group_by(Predictor) %>%
  do(plot = ggplot(., aes(x = y, y = yhat)) +
       geom_point(size = 0.5) +
       geom_abline(slope = 1, linetype = "dashed", color = "blue") +
       facet_wrap(~ Trait, nrow = 2, scales = "free") +
       theme(axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1)) +
       xlab("Observed") +
       ylab("Predicted") +
       ggtitle(.$Predictor))
invisible(lapply(full_y_yhat$plot, FUN = print))
```

>>>>

# LOOCV-Prediction
```{r message = FALSE, echo = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("tidyverse", "data.table", "dtplyr", "ggthemes", "viridis",
               "stringr", "forcats")

knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
                      warning = FALSE, echo = FALSE)
```

```{r Load-LOOCV-Data, cache = TRUE}
log_file <- fread("./data/derived/pred_log.txt")

pred_df <- log_file %>%
  .$Job_ID %>%
  as.list() %>%
  map(~readRDS(paste0("./data/derived/predictions/", ., ".RDS"))) %>%
  bind_rows(.id = "id") %>%
  left_join(y = log_file %>% select(Job_ID, Pred1, Pred2, Pred3, Trait, Iter),
            by = "Job_ID") %>%
  filter(Iter == 30000) %>%
  unite(col = Predictor, Pred1, Pred2, Pred3, sep = "-") %>%
  mutate(Trait = fct_recode(Trait,
    "DMY" = "GTM",
    "DMC" = "GTS",
    "FAT" = "FETT",
    "PRO" = "RPR",
    "SUG" = "XZ"
  )) %>%
  mutate(Trait = fct_relevel(Trait,
                             "DMY", "DMC", "ADF", "FAT", "PRO", "STA", "SUG"),
         Predictor = as.factor(Predictor)) %>%
  select(-Job_ID, -Iter)

# Update the variable 'Trait' for subsequent joins with 'pred_df'.
log_file <- log_file %>%
  filter(Iter == 30000) %>%
  mutate(Trait = fct_recode(Trait,
    "DMY" = "GTM",
    "DMC" = "GTS",
    "FAT" = "FETT",
    "PRO" = "RPR",
    "SUG" = "XZ"
  )) %>%
  mutate(Trait = fct_relevel(Trait,
                             "DMY", "DMC", "ADF", "FAT", "PRO",
                             "STA", "SUG")) %>%
  unite(col = Predictor, Pred1, Pred2, Pred3, sep = "-")
```

Compute the average standard deviation for each combination of predictors and 
traits -> http://stats.stackexchange.com/a/26647
```{r Coefficient-of-Variation, cache = TRUE}
coefficient_of_variation <- function(x, y) {
  sqrt(mean(x)) / mean(y)
}
cv_df <- pred_df %>%
  split(list(.$Trait, .$Predictor)) %>%
  map(~coefficient_of_variation(x = .$var_yhat, y = .$yhat)) %>%
  stack() %>%
  separate(col = ind, into = c("Trait", "Predictor"), sep = "[.]") %>%
  rename(CV = values) %>%
  tbl_dt() %>%
  mutate(Predictor = as.factor(Predictor),
         Trait = as.factor(Trait))
```

Combine predictive abilities and coefficients of variation for all predictor
and trait combinations for plotting.
```{r Organize-Data, cache = TRUE}
overview_df <- pred_df %>%
  group_by(Predictor, Trait) %>%
  summarize(r = cor(y, yhat)) %>%
  ungroup() %>%
  as.data.frame() %>%
  left_join(y = cv_df, by = c("Predictor", "Trait")) %>%
  as_tibble() %>%
  mutate(Predictor = as.factor(Predictor),
         Predictor = forcats::fct_reorder(Predictor, CV),
         Trait = fct_relevel(Trait,
                             "DMY", "DMC", "ADF", "FAT", "PRO",
                             "STA", "SUG"),
         Group = ifelse(Predictor %in% c("snp42-none-none",
                                         "ped42-none-none",
                                         "mrna42-none-none"),
                        yes = "Reduced", no = "Full")
  ) %>%
  mutate(Group = as.factor(Group))

# Specify a unified predictor order.
pred_order <- c("ped42-none-none", "snp42-none-none", "mrna42-none-none",
                "ped100-none-none", "snp100-none-none", "ped100-snp77-none",
                "ped100-mrna42-none", "snp100-mrna42-none",
                "ped100-snp77-mrna42")
```

Distribution of predicted and observed values.
Bear in mind that the distribution of the observed values will be different
for different predictors because some predictors "xx42" and "xx77", where
'xx' denotes a predictor, contain less hybrids than there are available 
(i.e. "xx100").
```{r Distribution-y-yhat, cache = TRUE}
legend_labs <- expression(y, hat(y))
distr_plot <- pred_df %>%
  filter(Predictor != "snp77-none-none") %>%
  gather(key = Value, value = `Phenotypic Value`, y, yhat) %>%
  mutate(Predictor = fct_relevel(Predictor, pred_order)) %>%
  ggplot(aes(x = `Phenotypic Value`, color = Value)) +
  geom_freqpoly() +
  scale_color_manual(values = viridis(length(legend_labs), option = "C"),
                     labels = legend_labs) +
  facet_grid(Predictor ~ Trait, scales = "free") +
  theme(legend.position = "top",
        strip.text.y = element_text(size = 5)) +
  ylab("Frequency")
```

Compare the coefficients of variation for all predictor-trait combinations.
```{r Coefficient-of-Variation-Plot, cache = TRUE}
cv_plot <- overview_df %>%
  filter(Predictor != "snp77-none-none") %>%
  mutate(Predictor = fct_relevel(Predictor, pred_order)) %>%
  ggplot(aes(x = CV, y = Predictor)) +
  geom_segment(aes(yend = Predictor), xend = 0, color = "gray50") +
  geom_point(aes(color = Predictor)) +
  scale_color_viridis(discrete = TRUE) +
  facet_wrap(~ Trait, scales = "free_x") +
  theme_bw() +
  theme(panel.grid.major.y = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = c(0.7, 0.13)
  ) +
  guides(color = guide_legend(ncol = 2,
                              reverse = FALSE)) +
  xlab("Coefficient of Variation") +
  ylab("Predictor Set") +
  xlim(0, NA)
```

Plot the predictive abilities for all predictor-trait combinations 
separately for the reduced and the full data set, respectively.
```{r Predictive-Ability-Plots, cache = TRUE}
pred_plot <- overview_df %>%
  as_tibble() %>%
  filter(Predictor != "snp77-none-none") %>%
  mutate(Predictor = fct_relevel(Predictor, pred_order),
         Group = fct_relevel(Group, "Reduced", "Full")
  ) %>%
  ggplot(aes(x = Predictor, y = r)) +
  geom_bar(stat = "identity", aes(fill = Predictor), color = "black") +
  geom_text(aes(label = round(r, digits = 3)), vjust = -0.2) +
  scale_fill_viridis(discrete = TRUE, option = "inferno") +
  facet_grid(Trait ~ Group, space = "free", scales = "free_x") +
  theme_bw() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    legend.position = "bottom",
    legend.key = element_rect(color = "white")
  ) +
  guides(fill = guide_legend(nrow = 3,
                             byrow = TRUE,
                             keywidth = 0.25,
                             keyheight = 0.25,
                             default.unit = "inch",
                             title.position = "top")
  ) +
  ylim(c(0, 1))
```
